---
title: "Gun Deaths in the United States, 2019"
author: "Carson Duce"
date: "r.system"
output: html_document
---
## Background
In a time of politically charged debates surrounding the topic of the 2nd Amendment to the US Constitution, both sides with their arguments often stretch and bend the present data to fit their message. Generalizing, the two sides we currently see are gun control advocates (GCA) and the 2nd Amendment Advocates (2AA). The GCA aim to reduce gun violence by enacting red flag laws, require a federal universal background check of which to ensure that only law abiding citizens in good mental health have access to firearms and its accessories. The 2AA aim to allow firearm ownership to be practiced by law abiding citizens as is stated by the 2nd Amendment to the United States Constitution. 

The 2nd Amendment reads: "A well regulated militia, being necessary to the security of a free State, the right of the people to keep and bear Arms, shall not be infringed." 

When looking at the history of the United States, the word choice of the 2nd Amendment, and the  and the wording chosen in accordance with its time, this phrase reads to the individual and what they are allowed as their rights. 

## Data

## Method

## Results

## Conclusion

## References
Willingham, AJ. “27 Words: Deconstructing the Second Amendment.” CNN, March 28, 2018. https://www.cnn.com/2016/08/10/politics/what-does-the-second-amendment-actually-mean-trnd/index.html. 

```{r}
library(readr)
library(dplyr)
library(tmaptools)
```


## Data Importation
```{r}

##data importation
handgun1 <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunArchiveDataDec2nd2019ThroughDec31st2019HandgunIncidents.csv")
rifle <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunArchiveDataJan1st2019ThroughDec31st2019RifleIncidents.csv")
handgun2 <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunArchiveDataJan1st2019ThroughMar1st2019HandgunIncidents.csv")
handgun3 <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunArchiveDataJul2nd2019ThroughSep1st2019HandgunIncidents.csv")
handgun4 <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunArchiveDataMar2nd2019ThroughMay1st2019HandgunIncidents.csv")
handgun5 <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunArchiveDataMay2nd2019ThroughJul1st2019HandgunIncidents.csv")
handgun6 <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunArchiveDataSep2nd2019ThroughDec1st2019HandgunIncidents.csv")
legalDefenseByState <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/usDataOnLegalityOfDefense.csv")
shotgun <- read_csv("https://raw.githubusercontent.com/Duceman56/PublicData/main/gunViolenceJan1stThroughDec31stShotgunIncidents.csv")
```

## select only wanted columns
```{r}
handgun1 <- handgun1[, 1:7]
handgun2 <- handgun2[, 1:7]
handgun3 <- handgun3[, 1:7]
handgun4 <- handgun4[, 1:7]
handgun5<- handgun5[, 1:7]
handgun6 <- handgun6[, 1:7]
shotgun <- shotgun[, 1:7]
```

## Combine the handgun data frames into one large data frame
```{r}
handgun <- rbind(handgun1, handgun2, handgun3, handgun4, handgun5, handgun6)
wiHandgun <- subset(handgun, handgun$State == "Wisconsin")
```

## combine the individual address, state/county, and state into one field
```{r}
wholeAddress <- paste(handgun$Address, handgun$`City Or County`, handgun$State, sep = ", ") ## for whole handgun data set
handgunTemp <- as.data.frame(wholeAddress) ## for whole handgun data set

wholeAddressWi <- paste(wiHandgun$Address, wiHandgun$`City Or County`, wiHandgun$State, sep = ", ") ## for wisconsin handgun data set
handgunWi <- as.data.frame(wholeAddressWi) ## for wisconsin handgun data set

wholeAddress1 <- paste(handgun1$Address, handgun1$`City Or County`, handgun1$State, sep = ", ")
handgunTemp1 <- as.data.frame(wholeAddress1)

wholeAddress2 <- paste(handgun2$Address, handgun2$`City Or County`, handgun2$State, sep = ", ")
handgunTemp2 <- as.data.frame(wholeAddress2)

wholeAddress3 <- paste(handgun3$Address, handgun3$`City Or County`, handgun3$State, sep = ", ")
handgunTemp3 <- as.data.frame(wholeAddress3)

wholeAddress4 <- paste(handgun4$Address, handgun4$`City Or County`, handgun4$State, sep = ", ")
handgunTemp4 <- as.data.frame(wholeAddress4)

wholeAddress5 <- paste(handgun5$Address, handgun5$`City Or County`, handgun5$State, sep = ", ")
handgunTemp5 <- as.data.frame(wholeAddress5)

wholeAddress6 <- paste(handgun6$Address, handgun6$`City Or County`, handgun6$State, sep = ", ")
handgunTemp6 <- as.data.frame(wholeAddress6)

wholeAddressRifle <- paste(rifle$Address, rifle$`City Or County`, rifle$State, sep = ", ")
rifleTemp1 <- as.data.frame(wholeAddressRifle)

wholeAddressShotgun <- paste(shotgun$Address, shotgun$`City Or County`, shotgun$State, sep = ", ")
shotgunTemp1 <- as.data.frame(wholeAddressShotgun)
```

## add address to main handgun, rifles, and shotgun dataframe
```{r}
handgun <- cbind(handgun, handgunTemp)
rifle <- cbind(rifle, rifleTemp1)
shotgun <- cbind(shotgun, shotgunTemp1)
```

## geocode portion
```{r}
## geocoded <- geocode_OSM(handgunTemp$wholeAddress)
## geocodedWhole <- geocode_OSM(handgun$wholeAddress)
```

## summary statistics
```{r}
sum(handgun$`# Killed`)
sum(handgun$`# Injured`)
str(handgun)

sum(rifle$`# Killed`)
sum(rifle$`# Injured`)
str(rifle)

sum(shotgun$`# Victims Killed`)
sum(shotgun$`# Victims Injured`)
str(shotgun)
```


## write new csv to dedicated locations
```{r}
write_csv(handgun, file = "//Users//carsond//Desktop//handgun.csv") #"Q://StudentCoursework//Zeitler//GEOG.280.001.2235//DUCECR1126//09 - Final Map Project//handgun.csv")
write_csv(rifle, file = "//Users//carsond//Desktop//rifle.csv")
write_csv(shotgun, file = "//Users//carsond//Desktop//shotgun.csv")
```
